Galaxy Zoo: Morphological Classifications for 150,000 Galaxies in HST Legacy Imaging

### 2016-05-20 writeup/gz_hubble_data.tex
edits to text in response to comments in issue 140

The approach used in GZH for correcting the weighted classifications for
redshift bias rests on the assumption that the \emph{degree} of bias is a
function of the apparent size and brightness of the image as seen on screen.
This is controlled by two types of parameters: \textbf{intrinsic} properties of
the galaxy itself, such as its physical diameter and luminosity, and
\textbf{extrinsic} properties, such as the distance (redshift) of the galaxy
and its relative orientation. There may also be other parameters that affect
classifier accuracy, such as the proximity of close companions or 
individual classifier ability and skill. The combination of all
such parameters forms a high-dimensional space, and it is not clear how to
separate this into individual effects. Instead, the method used here employs
only two parameters intended to capture the bulk of the change in bias: the
$r^\prime$-band surface brightness ($\mu_r$; intrinsic) and redshift ($z$;
extrinsic).

This dependence of the vote fraction on $\mu_r$ and \zsim{} is measured using the
\ferengi{} images over all the evolutionary correction factors. It is assumed
that the ``true'' (ie, debiased) vote fraction $f_{\mu,z}$ for an
artificially-redshifted galaxy can be expressed as:

@@

\noindent where $f_{\mu,z=0.3}$ is the ``calibrated'' vote fraction at the
lowest redshift in the \ferengi{} bins ($z=0.3$) and $\zeta$ is a positive
parameter that controls the rate at which $f$ decreases with increasing
redshift. This formula fits the data relatively well (with very few exceptions,
the vote fractions for featured galaxies decrease monotonically with increasing
redshift), and the exponential function bounds the observed vote fractions
between $f_{\mu,z=0.3}$ and zero. Figure~\ref{fig:zeta_examples} shows the
change in vote fraction and the fit to Equation~\ref{eqn:fzeta} for a random
selection of galaxies in the \ferengi{} images.

@@

\noindent where $\hat\zeta$ is the correction factor applied to each galaxy as
a function of surface brightness. The best-fit parameters to the linear fit
(from least-squares optimization) are $\zeta_0=0.1$, $\zeta_1=1.4$. To make the
final debiased correction, the simple exponential form of
Equation~\ref{eqn:fzeta} is modified to bound the debiased vote fractions
between $f$ and 1:

@@

The criteria for determining whether a region of this space is single-valued,
and therefore correctable, is as follows: In each surface brightness and
redshift bin, the relationship between \ffeaturesz{} and \ffeaturesrest{} is
modeled by fitting the data with a polynomial of degrees $n=3$, 2, and 1, and
using the best formal fit out of the three as measured by the sum of the
residuals. These fits are shown as the dashed black lines in
Figure~\ref{fig:f_vs_f}. Any flat regions of the polynomial fits are areas in which
there is not a clear single-valued relationship between \ffeaturesz{} and
\ffeaturesrest; this is quantified by setting a minimum slope cut of 0.4. Any
data in which the polynomial fit has a slope less than this value is considered
\emph{not} one-to-one, and therefore \ffeaturesz{} cannot be boosted to its
\ffeaturesrest{} value.  Galaxies in this region are referred to as the
\emph{lower limit} sample, because the most stringent correction available is
that the weighted \ffeatures{} is a lower limit to the true value.  These
regions are highlighted in blue in Figure~\ref{fig:f_vs_f}. Uncolored (white)
regions of the plot have sufficiently high slopes to consider the relationship
as single-valued; galaxies in these regions are considered ``correctable'', and
only these are used in measuring the parameters for the $\zeta$ function
(Section~\ref{ssec:zeta}). Only surface brightness/redshift bins with at least
5 galaxies were considered; regions with fewer than 5 galaxies are considered
to have ``not enough information'' to determine the \ffeaturesz{} and
\ffeaturesrest{} relationship, colored gray in Figure~\ref{fig:f_vs_f}.

>>>>

The approach used in GZH for correcting the weighted classifications for
redshift bias rests on the assumption that the \emph{degree} of bias is a
function of the apparent size and brightness of the image as seen on screen.
This is controlled by two types of parameters: \textbf{intrinsic} properties of
the galaxy itself, such as its physical diameter and luminosity, and
\textbf{extrinsic} properties, such as the distance (redshift) of the galaxy
and its relative orientation. There may also be other parameters that affect
classifier accuracy, such as the proximity of close companions or 
individual classifier ability and skill. The combination of all
such parameters forms a high-dimensional space, and it is not clear how to
separate this into individual effects. Instead, the method used here employs
only two parameters intended to capture the bulk of the change in bias: the
$i$-band surface brightness ($\mu_i$; intrinsic) and redshift ($z$;
extrinsic).

This dependence of the vote fraction on $\mu_i$ and \zsim{} is measured using the
\ferengi{} images over all the evolutionary correction factors. It is assumed
that the ``true'' (ie, debiased) vote fraction $f_{\mu,z}$ for an
artificially-redshifted galaxy can be expressed as:

@@

\noindent where $f_{\mu,z=0.3}$ is the ``calibrated'' vote fraction at the
lowest redshift in the \ferengi{} bins ($z=0.3$) and $\zeta$ is a positive
parameter that controls the rate at which $f$ decreases with increasing
redshift. The exponential function bounds the observed vote fractions
between $f_{\mu,z=0.3}$ and zero based on two assumptions: 1) the vote fractions for featured galaxies decrease monotonically with increasing redshift, therefore restricting vote fractions for a given galaxy to be less than $f_{\mu,z=0.3}$  (which is true of the data with very few exceptions), and 2) the vote fractions cannot be less than zero (which is always true). Figure~\ref{fig:zeta_examples} shows the
change in vote fraction and the fit to Equation~\ref{eqn:fzeta} for a random
selection of galaxies in the \ferengi{} images.

@@

\noindent where $\hat\zeta$ is the correction factor applied to each galaxy as
a function of surface brightness. The best-fit parameters to the linear fit
(from least-squares optimization) are $\zeta_0=0.75$, $\zeta_1=-.04$. To make the
final debiased correction, the simple exponential form of
Equation~\ref{eqn:fzeta} is modified to bound the debiased vote fractions
between $f$ and 1:

@@

The criteria for determining whether a region of this space is single-valued,
and therefore correctable, is as follows: In each surface brightness and
redshift bin, the relationship between \ffeaturesz{} and \ffeaturesrest{} is
modeled by fitting the data with a polynomial of degrees $n=3$, 2, and 1, and
using the best formal fit out of the three as measured by the sum of the
residuals. These fits are shown as the dashed black lines in
Figure~\ref{fig:f_vs_f}. Any flat regions of the polynomial fits are areas in which
there is not a clear single-valued relationship between \ffeaturesz{} and
\ffeaturesrest; this is quantified by setting a minimum slope cut of 0.4. Any
data in which the polynomial fit has a slope less than this value is considered
\emph{not} one-to-one, and therefore \ffeaturesz{} cannot be boosted to its
\ffeaturesrest{} value.  Galaxies in this region are referred to as the
\emph{lower limit} sample, because the most stringent correction available is
that the weighted \ffeatures{} is a lower limit to the true value.  These
regions are highlighted in blue in Figure~\ref{fig:f_vs_f}. Uncolored (white)
regions of the plot have sufficiently high slopes to consider the relationship
as single-valued; galaxies in these regions are considered ``correctable'', and
only these are used in measuring the parameters for the $\zeta$ function
(Section~\ref{ssec:zeta}). Only surface brightness/redshift bins with at least
5 galaxies were considered; regions with fewer than 5 galaxies are considered
to have ``not enough information'' to determine the \ffeaturesz{} and
\ffeaturesrest{} relationship, colored gray in Figure~\ref{fig:f_vs_f}. These galaxies are hereafter referred to as the ``NEI'' sample.
